---
title: "MOGTree"
output:
  github_document:
    toc: false
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# MOGTree

`MOGTree` implements the multi-objective gain tree (MOG-Tree) workflow for
learning patient subgroups with distinct treatment-outcome contrast profiles.
The package centers on four core steps:

1. fit a tree partition with `DTRtree()`;
2. assign new subjects to terminal nodes with `predict_leaf.DTR()`;
3. estimate propensity scores with `M.propen()` and outcome regressions with
   `Reg.mu()`;
4. summarize leaf-specific treatment preferences across a grid of outcome
   weights with `mus.AIPW()`.

The current implementation matches the functions used in the repository's
simulation studies. In particular, the tree fit returns the subgroup structure,
and the weight-specific treatment summaries are built from the AIPW estimates.

## Installation

```r
# install.packages("remotes")
remotes::install_github("SelinaSong0412/MOGTree")
```

For local development from the repository root:

```r
devtools::install(".")
```

## Example workflow

```{r minimal-workflow, eval = FALSE}
library(MOGTree)

set.seed(1)
n <- 120
X1 <- rnorm(n)
X2 <- rnorm(n)
H <- data.frame(X1 = X1, X2 = X2)
A <- sample(1:4, n, replace = TRUE)

group <- ifelse(X1 <= 0, 1, 2)
base_y1 <- rbind(
  c(1.2, 0.4, -0.1, -0.4),
  c(-0.2, 0.3, 0.8, 1.1)
)
base_y2 <- rbind(
  c(0.1, 0.7, 1.0, 0.6),
  c(1.1, 0.8, 0.2, -0.1)
)

Y1_cf <- matrix(NA_real_, n, 4)
Y2_cf <- matrix(NA_real_, n, 4)
for (i in seq_len(n)) {
  Y1_cf[i, ] <- base_y1[group[i], ] + 0.2 * X2[i] + rnorm(4, sd = 0.15)
  Y2_cf[i, ] <- base_y2[group[i], ] - 0.1 * X2[i] + rnorm(4, sd = 0.15)
}

Y1_obs <- Y1_cf[cbind(seq_len(n), A)]
Y2_obs <- Y2_cf[cbind(seq_len(n), A)]

w <- seq(0, 1, by = 0.25)
tree_fit <- DTRtree(
  Ys = cbind(Y1_obs, Y2_obs),
  A = A,
  H = H,
  depth = 2,
  minsplit = 15,
  w_vec = cbind(w, 1 - w),
  weight_combine = "mean",
  lambda = 0.01
)

leaf_id <- predict_leaf.DTR(tree_fit, H)
table(leaf_id)

pi_hat <- M.propen(A, H)
mu1_reg <- Reg.mu(Y1_obs, A, H)$mus.reg
mu2_reg <- Reg.mu(Y2_obs, A, H)$mus.reg
mu1_hat <- mus.AIPW(Y1_obs, A, pi_hat, mu1_reg)
mu2_hat <- mus.AIPW(Y2_obs, A, pi_hat, mu2_reg)

one_leaf <- sort(unique(leaf_id))[1]
idx <- leaf_id == one_leaf

sapply(w, function(weight) {
  weighted_mean <- colMeans(weight * mu1_hat[idx, ] + (1 - weight) * mu2_hat[idx, ])
  which.max(weighted_mean)
})
```

```{r}
tree_fit
```


## Main functions

- `DTRtree()` builds the MOG-Tree partition using a weight-aggregated gain
  criterion.
- `predict_leaf.DTR()` maps new covariate profiles to terminal nodes.
- `M.propen()` estimates multinomial propensity scores.
- `Reg.mu()` estimates treatment-specific outcome regressions.
- `mus.AIPW()` combines the propensity and regression models into an
  augmented inverse probability weighted estimator.

## Learn more

The package vignette walks through the same workflow with a small reproducible
example:

- `vignette("example", package = "MOGTree")`
