---
title: "Getting Started with MOGTree"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with MOGTree}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The manuscript for MOG-Tree motivates a tree that is stable across a grid of
outcome weights by maximizing an aggregated multi-objective gain. In the
current repository implementation, the package workflow has two layers:

1. `DTRtree()` learns a subgrouping tree from observed outcomes, treatments,
   and baseline covariates.
2. `M.propen()`, `Reg.mu()`, and `mus.AIPW()` can then be used to summarize
   which treatment is preferred inside each leaf for different outcome weights.

This vignette shows that workflow with a small two-outcome toy example.

## Simulate a small example

```{r}
library(MOGTree)

set.seed(412)
n <- 120
H <- data.frame(
  X1 = rnorm(n),
  X2 = rnorm(n)
)
A <- sample(1:4, n, replace = TRUE)

group <- ifelse(H$X1 <= 0, 1, 2)
base_y1 <- rbind(
  c(1.2, 0.4, -0.1, -0.4),
  c(-0.2, 0.3, 0.8, 1.1)
)
base_y2 <- rbind(
  c(0.1, 0.7, 1.0, 0.6),
  c(1.1, 0.8, 0.2, -0.1)
)

Y1_cf <- matrix(NA_real_, n, 4)
Y2_cf <- matrix(NA_real_, n, 4)

for (i in seq_len(n)) {
  Y1_cf[i, ] <- base_y1[group[i], ] + 0.2 * H$X2[i] + rnorm(4, sd = 0.15)
  Y2_cf[i, ] <- base_y2[group[i], ] - 0.1 * H$X2[i] + rnorm(4, sd = 0.15)
}

Y1_obs <- Y1_cf[cbind(seq_len(n), A)]
Y2_obs <- Y2_cf[cbind(seq_len(n), A)]
```

## Fit the MOG-Tree partition

```{r}
weight_grid <- seq(0, 1, by = 0.25)

tree_fit <- DTRtree(
  Ys = cbind(Y1_obs, Y2_obs),
  A = A,
  H = H,
  depth = 2,
  minsplit = 15,
  w_vec = cbind(weight_grid, 1 - weight_grid),
  weight_combine = "mean",
  lambda = 0.01
)

tree_fit
```

The fitted object records the split variable, split cutoff, and terminal-group
index for each node.

## Assign subjects to terminal nodes

```{r}
leaf_id <- predict_leaf.DTR(tree_fit, H)
table(leaf_id)
```

## Estimate leaf-specific treatment preferences

The tree partitions patients into subgroups. To study treatment preferences
inside each subgroup, we estimate propensity scores, regression means, and AIPW
counterfactual means for each outcome.

```{r}
pi_hat <- M.propen(A, H)
mu1_reg <- Reg.mu(Y1_obs, A, H)$mus.reg
mu2_reg <- Reg.mu(Y2_obs, A, H)$mus.reg
mu1_hat <- mus.AIPW(Y1_obs, A, pi_hat, mu1_reg)
mu2_hat <- mus.AIPW(Y2_obs, A, pi_hat, mu2_reg)
```

The code below computes, for each terminal node and each outcome weight, the
mean weighted counterfactual outcome for every treatment and reports the
treatment with the largest value.

```{r}
summarize_leaf_preferences <- function(leaf_membership, mu1_hat, mu2_hat, weights) {
  leaf_values <- sort(unique(leaf_membership))

  do.call(
    rbind,
    lapply(leaf_values, function(leaf) {
      idx <- leaf_membership == leaf
      do.call(
        rbind,
        lapply(weights, function(weight) {
          weighted_mean <- colMeans(weight * mu1_hat[idx, ] + (1 - weight) * mu2_hat[idx, ])
          data.frame(
            leaf = leaf,
            weight_outcome1 = weight,
            preferred_treatment = which.max(weighted_mean)
          )
        })
      )
    })
  )
}

preference_summary <- summarize_leaf_preferences(
  leaf_membership = leaf_id,
  mu1_hat = mu1_hat,
  mu2_hat = mu2_hat,
  weights = weight_grid
)

preference_summary
```

This summary shows how the preferred treatment can vary across terminal nodes and across different weights on the first outcome.

## Input checklist

For your own data:

- `Ys` should be an `n x m` matrix of observed outcomes.
- `A` should be a treatment vector coded with at least two observed treatment levels.
- `H` should contain the baseline tailoring variables used to split the tree.
- `w_vec` should be a matrix whose rows are candidate weight vectors that sum
  to one.

The simulation scripts in the repository's
`Simulation/Two_Outcome_Simulation.Rmd` and
`Simulation/Three_Outcome_Simulation.Rmd` files provide larger end-to-end examples matching the experiments in the manuscript.
