---
title: "Three-Outcome Simulation Code"
output: html_document
date: '2024-04-04'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggpubr)
library(ggplot2)
library(gridExtra)
library(cowplot)
library(grid)
library(DiagrammeR)
library(nnet)
# for performance comparison
library(mclust)
library(kernlab)
library(kohonen)
library(class)
library(mclust)
library(clue) # NMI
library(cluster)
library(fpc) # RI
library(tibble)
library(dplyr)
library(tidyr)
# Load necessary libraries
library(plotly)


library(MOTRL)

setwd("R")
source("Benefit_gain.R")
source("best.H.R")
source("MO_Tree.R")
source("predict_leaf.R")
source("Reg.mu.R")
source("Split.X.R")
source("visualization.R")
source("vis_tree.R")

source("mus.AIPW.R")
source("M.propen.R")
Custom.color = c("#CD2626","#79aec1","#2E8B57","#d6b55a")

map_to_consecutive = function(x) {
  mapped_values = match(x, unique(x))
  return(mapped_values)
}
```


## YS update simulation function with new designed pareto dominances

```{r}
Simulation_MOTree_3 = function(iter = 100, n.cov = 5, n.per.group = 100) {
  
  nmi_results = rep(list(rep(NA, iter)),7) 
  ari_results = rep(list(rep(NA, iter)),7) 
  weights <- expand.grid(seq(0, 1, by = 0.1), seq(0, 1, by = 0.1)) %>%
    filter(Var1 + Var2 <= 1) %>%
    mutate(Var3 = round((1 - Var1 - Var2), 1))
  
  for (w_idx in 1:nrow(weights)) {
    assign(paste0("avg_cf_results", w_idx), rep(list(rep(NA, iter)),7) )
    assign(paste0("true_optimal_weighted_mean", w_idx), numeric(iter))  
    assign(paste0("non_dominant_results", w_idx), rep(list(rep(NA, iter)), 7))  
    
  }
  
  for (i in c(1:iter)) {
    
    ###########################################
    ############ Data generation ##############
    ###########################################
    set.seed(412 + i)
    k = 4 # number of treatment
    n_grp = 8  # Number of groups
    total_patients <- n.per.group * n_grp
        
    # group = rep(1:4, each = n.per.group)
    groups <- 1:n_grp   # Total groups
    group <- rep(groups, each = n.per.group)

    
    X <- matrix(NA, nrow = n.per.group * n_grp, ncol = n.cov)
    X[, 1] <- rnorm(n.per.group * n_grp, 0, 0.05) + (group - 1) / 5
    X[, 2] <- rnorm(n.per.group * n_grp, mean = 0, sd = 0.05) - 0.5 * (group - 4)
    for (jj in 3:n.cov) {
      X[, jj] <- rnorm(n.per.group * n_grp, mean = 0, sd = 0.05)
    }
    
    # m1 <- c(4, 1, 0.5, 0, 3, 0, 0.5, 0, 4, 0, 0.5, 1)   # Treatment 1 dominates for Group 1 and 2
    # m2 <- c(4, 0, 0, 1.5, 3, 0.5, 0, 0.5, 4, 1, 0, 0.5)   # Same dominance structure as Group 1
    # m3 <- c(4, 3, 2, 1, 3, 2, 1, 0, 4, 3, 1, 1)   # Treatments 1 and 2 dominate
    # m4 <- c(3, 3, 3, 2, 4, 4, 4, 2, 3, 3, 2, 1)   # Treatments 2 and 3 dominate
    # m5 <- c(1, 0.5, 0, 4, 0, 0, 1, 4, 0.5, 0, 1, 4)   # Treatment 4 dominates for Group 5 and 6
    # m6 <- c(1.5, 1, 1.5, 4, 1, 1.5, 0, 4, 0, 1, 0, 4)   # Same dominance structure as Group 5
    # m7 <- c(0, 0, 4, 4, 1, 1, 4, 4, 0, 0, 4, 5)   # Treatments 3 and 4 dominate
    # m8 <- c(0, 4, 0, 4, 1, 5, 1, 4, 0, 4, 0, 4)   # Treatments 2 and 4 dominate
    m1 <- c(4, 1, 0.5, 0, 3, 0, 1.5, 0, 4, 0, 0.5, 1)   # Treatment 1 dominates for Group 1 and 2
    m2 <- c(4, 0, 0, 1.5, 3, 0.5, 0, 0.5, 4, 1, 0, 0.5)   # Same dominance structure as Group 1
    m3 <- c(3, 1.5, 0.5, 0, 4, 0, 0, 1, 4, 1, 2, 0.5)    # Treatments 1 and 2 dominate
    m4 <- c(4, 3, 2, 1, 1, 4, 1, 4, 2.5, 3, 0, 1)   # Treatments 2 and 3 dominate
    m5 <- c(1, 0.5, 0, 4, 2, 0, 1, 4, 0.5, 0, 1, 4)   # Treatment 4 dominates for Group 5 and 6
    m6 <- c(1.5, 1, 0.5, 4, 1, 1.5, 0, 4, 2, 1, 0, 4)   # Same dominance structure as Group 5
    m7 <- c(0, 0, 4, 2.5, 0, 1, 3, 2, 0, 0, 4, 1)   # Treatments 3 and 4 dominate
    m8 <- c(0, 4, 0, 2, 0, 1, 0, 2, 0.5, 2.5, 3, 4)   # Treatments 2 and 4 dominate
    list_m <- list(m1, m2, m3, m4, m5, m6, m7, m8)
    
    Y <- matrix(0, n.per.group * n_grp, 3 * k)
    for (ii in 1:(n.per.group * n_grp)) {
      ind <- group[ii]
      Y[ii, ] <- list_m[[ind]] +
        X[ii, 1] * 0.1 +
        X[ii, 2] * 0.05 +
        X[ii, 3] * -0.0005 +
        X[ii, 4] * 0.0001 +
        X[ii, 5] * -0.0001 +
        rnorm(12, 0, 0.3)
    }
    Y1 <- Y[, c(1, 4, 7, 10)]
    Y2 <- Y[, c(2, 5, 8, 11)]
    Y3 <- Y[, c(3, 6, 9, 12)]
    
    true_treatment <- sample(1:k, n.per.group * n_grp, replace = TRUE)
    Y1.obs <- sapply(1:(n.per.group * n_grp), function(i) Y1[i, true_treatment[i]])
    Y2.obs <- sapply(1:(n.per.group * n_grp), function(i) Y2[i, true_treatment[i]])
    Y3.obs <- sapply(1:(n.per.group * n_grp), function(i) Y3[i, true_treatment[i]])
    
    df_obs <- data.frame(
      Y1_obs = Y1.obs,
      Y2_obs = Y2.obs,
      Y3_obs = Y3.obs,
      Treat_num = true_treatment,
      Group = factor(group)
    )
    df_obs$true_Group <- case_when(
      df_obs$Group %in% c(1, 2, 3) ~ 1,
      df_obs$Group %in% c(4) ~ 2,
      df_obs$Group %in% c(5, 6) ~ 3,
      df_obs$Group %in% c(7) ~ 4,
      df_obs$Group %in% c(8) ~ 5
    )
    df_obs$Treatment <- as.factor(case_when(
      df_obs$Treat_num == 1 ~ "A",
      df_obs$Treat_num == 2 ~ "B",
      df_obs$Treat_num == 3 ~ "C",
      df_obs$Treat_num == 4 ~ "D"
    ))
    # for (cov in 1:j) {
    for (cov in 1:n.cov) {
      df_obs[[paste0("X", cov)]] <- X[, cov]
    }
    # X <- matrix(NA, nrow = n.per.group * n_grp, ncol = n.cov)
    # X[, 1] <- rnorm(n.per.group * n_grp, 0, 0.05) + (group - 1) / 3
    # X[, 2] <- rnorm(n.per.group * n_grp, mean = 0, sd = 0.05) - 0.5 * (group - 2)
    # for (k in 3:n.cov) {
    #   X[, k] <- rnorm(n.per.group * n_grp, mean = 0, sd = 0.05)
    # }
    # 
    # n_Y <- 3  # Number of outcomes (Y1, Y2, Y3)
    # Parameter <- matrix(rnorm(n_grp * k * n_Y, 0, 1), nrow = n_grp, ncol = k * n_Y)
    # Y <- matrix(0, nrow = n.per.group * n_grp, ncol = k * n_Y)
    # for (ii in 1:(n.per.group * n_grp)) {
    #   ind <- group[ii]
    #   Y[ii, ] <- Parameter[ind, ] + 
    #     rnorm(k * n_Y, 0, 0.2)
    # }
    # Y_record <- lapply(1:n_Y, function(yy) Y[, ((yy - 1) * k + 1):(yy * k)])
    # Y1 <- Y_record[[1]]
    # Y2 <- Y_record[[2]]
    # Y3 <- Y_record[[3]]
    # 
    # 
    # true_treatment <- sample(1:k, n.per.group * n_grp, replace = TRUE)
    # Y1.obs <- sapply(1:(n.per.group * n_grp), function(i) Y1[i, true_treatment[i]])
    # Y2.obs <- sapply(1:(n.per.group * n_grp), function(i) Y2[i, true_treatment[i]])
    # Y3.obs <- sapply(1:(n.per.group * n_grp), function(i) Y3[i, true_treatment[i]])
    # 
    # df_obs <- data.frame(
    #   Y1_obs = Y1.obs,
    #   Y2_obs = Y2.obs,
    #   Y3_obs = Y3.obs,
    #   Treat_num = true_treatment,
    #   Group = factor(group)
    # )
    # 
    # # df_obs$Treatment <- factor(
    # #   df_obs$Treat_num,
    # #   labels = c("A", "B", "C", "D")
    # # )
    # df_obs$Treatment = as.factor(case_when(df_obs$Treat_num == 1 ~ "A",
    #                                        df_obs$Treat_num == 2 ~ "B",
    #                                        df_obs$Treat_num == 3 ~ "C",
    #                                        df_obs$Treat_num == 4 ~ "D"))
    # 
    # for (cov in 1:n.cov) {
    #   df_obs[[paste0("X", cov)]] <- X[, cov]
    # }
    
    ####### Pareto Dominance Evaluation #######
    non_dominant_treatments <- lapply(1:total_patients, function(patient) {
      outcomes <- cbind(Y1[patient, ], Y2[patient, ], Y3[patient, ])  
      is_non_dominated <- function(outcomes) {
        apply(outcomes, 1, function(x) {
          !any(apply(outcomes, 1, function(y) all(y >= x) & any(y > x)))
        })
      }
      pareto_front <- which(is_non_dominated(outcomes))
      return(pareto_front)  
    })
    
    ## true optimal weighted conterfactual mean
    for (w_idx in 1:nrow(weights)) {
      w <- as.numeric(weights[w_idx, ])
      
      # Weighted counterfactual outcomes for all treatments
      Y_w <- w[1] * Y1 + w[2] * Y2 + w[3] * Y3
      # Maximum weighted counterfactual outcome for each individual (Y*_opt)
      max_weighted_outcomes <- apply(Y_w, 1, max)
      assign(paste0("true_optimal_weighted_mean", w_idx),
             replace(get(paste0("true_optimal_weighted_mean", w_idx)), i, mean(max_weighted_outcomes)))
    }
    
    
    
    ###########################################
    ########## Clustering methods #############
    ###########################################
    
    ## MOTree
    Cov = df_obs %>% dplyr::select(starts_with("X"))
    treeout = DTRtree(Ys = cbind(df_obs$Y1_obs, df_obs$Y2_obs, df_obs$Y3_obs),
                      A = df_obs$Treat_num,
                      H = Cov,
                      pis.hat=NULL,
                      mus.reg=NULL,
                      depth=3,
                      minsplit=20,
                      w_vec = as.matrix(weights),
                      weight_combine = "mean",
                      lambda = 0.05)
    leaf_pred = predict_leaf.DTR(treeout, newdata = Cov)
    df_obs$group.MOTree = map_to_consecutive(leaf_pred)
    
    SimuData = df_obs %>% dplyr::select(# Y1_obs, Y2_obs, 
      # Treat_num, 
      starts_with("X"))
    
    ## K-means 
    kmeans_result5 = kmeans(SimuData, centers = 5)  # Assuming 5 clusters
    kmeans_result8 = kmeans(SimuData, centers = 8)  # Assuming 8 clusters
    df_obs$group.kmeans5 = kmeans_result5$cluster
    df_obs$group.kmeans8 = kmeans_result8$cluster
    
    ## Hierarchical
    hc_result = hclust(dist(SimuData), method = "ward.D") # "ward.D2")
    hc_clusters = cutree(hc_result, k = 8) # k = 4)
    df_obs$group.hc = hc_clusters
    
    ## GMM
    gmm_result = Mclust(SimuData) 
    df_obs$group.gmm = gmm_result$classification
    
    ## Spectral Clustering
    spectral_result = specc(as.matrix(SimuData), centers = 8)  # Assuming 8 clusters
    df_obs$group.spectral = spectral_result@.Data
    
    ## Self-Organizing Map (SOM)
    som_grid = kohonen::somgrid(xdim = 2, ydim = 4, topo = "rectangular")
    som_model = kohonen::som(as.matrix(SimuData), grid = som_grid, rlen = 100)
    som_clusters = cutree(hclust(dist(som_model$codes[[1]])), k = 8)  
    som_result = som_model$unit.classif
    df_obs$group.som = som_clusters[som_result]
    
    ###########################################
    ############ Begin comparison #############
    ###########################################
    true_labels_nmi = as.cl_partition(df_obs$true_Group)     
    true_labels_ari = as.numeric(df_obs$true_Group)      
    clustering_methods = list(
      MOTree = df_obs$group.MOTree,
      # MOTRL = NA,
      HC = df_obs$group.hc,
      GMM = df_obs$group.gmm,
      KMeans5 = df_obs$group.kmeans5,
      KMeans8 = df_obs$group.kmeans8,
      Spectral = df_obs$group.spectral,
      SOM = df_obs$group.som
    )
    
    dist_matrix = dist(as.matrix(SimuData))
    
    for (method in seq_along(clustering_methods)) {
      method_name = names(clustering_methods)[method]
      pred_labels_nmi = as.cl_partition(clustering_methods[[method_name]])  
      pred_labels_ari = as.numeric(clustering_methods[[method_name]])    
      cluster_labels = as.numeric(clustering_methods[[method_name]])      
      nmi_results[[method]][i] = as.numeric(cl_agreement(true_labels_nmi, 
                                                         pred_labels_nmi, 
                                                         method = "NMI"))
      ari_results[[method]][i] = cluster.stats(dist_matrix, true_labels_ari, 
                                               pred_labels_ari)$corrected.rand
      
      cluster_opt_trt = list()
      for (w_idx in 1:nrow(weights)) {
        w <- as.numeric(weights[w_idx, ])
        Y_w <- w[1] * Y1 + w[2] * Y2 + w[3] * Y3
        
        # find cluster-based optimal treatment for each individual (by each of the methods)
        cluster_opt_trt[[method]] = sapply(unique(cluster_labels), 
                                           function(cluster) {
                                             cluster_ids = which(cluster_labels == cluster)
                                             avg_outcomes = colMeans(Y_w[cluster_ids, , drop = FALSE])
                                             which.max(avg_outcomes)  
                                           })
        
        df_obs[[paste0("opt_trt_", method_name, "_w", w_idx)]] = sapply(cluster_labels,
                                                                        function(cluster) cluster_opt_trt[[method]][cluster]) 
        df_obs[[paste0("cf_", method_name, "_w", w_idx)]] = sapply(1:nrow(Y_w), 
                                                                   function(x) Y_w[x, df_obs[[paste0("opt_trt_", method_name, "_w", w_idx)]][x]])
        
        avg_cf_results <- get(paste0("avg_cf_results", w_idx))
        avg_cf_results[[method]][i] <- mean(df_obs[[paste0("cf_", method_name, "_w", w_idx)]], na.rm = T)
        assign(paste0("avg_cf_results", w_idx), avg_cf_results)
        
        
        
        ## YS ADD THIS NEW FOR NEW MATRIC 11/25/2024
        ## Predicted optimal treatment for current weight
        pred_opt_trt <- df_obs[[paste0("opt_trt_", method_name, "_w", w_idx)]]
        is_non_dominant <- sapply(seq_len(n.per.group * 4), function(patient) {
          pred_opt_trt[patient] %in% non_dominant_treatments[[patient]]
        })
        non_dominant_results <- get(paste0("non_dominant_results", w_idx))
        non_dominant_results[[method]][i] <- mean(is_non_dominant, na.rm = T)
        assign(paste0("non_dominant_results", w_idx), non_dominant_results)
        
        cluster_opt_trt = list() 
        
      }
      
    }
    
    
    
  }
  
  # evaluation
  nmi_mean = round(as.numeric(lapply(nmi_results, mean)),3)
  nmi_sd = round(as.numeric(lapply(nmi_results, sd)),3)
  ari_mean = round(as.numeric(lapply(ari_results, mean)),3)
  ari_sd = round(as.numeric(lapply(ari_results, sd)),3)
  nmi_out = paste0(nmi_mean, "(", nmi_sd, ")")
  ari_out = paste0(ari_mean, "(", ari_sd, ")")
  
  # non_dominant_summary <- matrix(NA, nrow = length(clustering_methods), ncol = nrow(weights))
  # rownames(non_dominant_summary) <- names(clustering_methods)
  #   # paste0("Method_", 1:7)  # Adjust according to actual method names
  # colnames(non_dominant_summary) <- paste0("Weight_", 1:nrow(weights))
  
  

  # weight specific results
  for (w_idx in 1:nrow(weights)) {
    
    avg_cf_results <- get(paste0("avg_cf_results", w_idx))
    assign(paste0("avg_cf_mean", w_idx), 
           round(as.numeric(lapply(avg_cf_results, mean)), 3))
    assign(paste0("avg_cf_sd", w_idx), 
           round(as.numeric(lapply(avg_cf_results, sd)), 3))
    assign(paste0("avg_cf_out", w_idx), 
           paste0(get(paste0("avg_cf_mean", w_idx)), 
                  "(", get(paste0("avg_cf_sd", w_idx)), ")"))
    
    non_dominant_means <- get(paste0("non_dominant_results", w_idx)) # [[method]]
    # non_dominant_summary[method, w_idx] <- mean(non_dominant_means)
    assign(paste0("non_dominant_mean", w_idx), 
           round(as.numeric(lapply(non_dominant_means, mean)), 3))
    assign(paste0("non_dominant_sd", w_idx), 
           round(as.numeric(lapply(non_dominant_means, sd)), 3))
    assign(paste0("non_dominant_out", w_idx), 
           paste0(get(paste0("non_dominant_mean", w_idx)), 
                  "(", get(paste0("non_dominant_sd", w_idx)), ")"))
  }
  
  
  true_optimal_weighted_means <- sapply(1:nrow(weights), function(w_idx) {
    mean(get(paste0("true_optimal_weighted_mean", w_idx)))
  })
  true_optimal_weighted_means <- t(data.frame(true_optimal_weighted_means))
  rownames(true_optimal_weighted_means) <- "True_Optimal_Mean"
  colnames(true_optimal_weighted_means) <- paste0("Weight_", 1:nrow(weights))# seq(0, 1, by = 0.1))
  
  
  clust_perform = data.frame(NMI = nmi_out,
                             ARI = ari_out)
  row.names(clust_perform) = names(clustering_methods)
  
  # cf_Y_perform = data.frame(
  #   avg_cf_w1 = avg_cf_out1,
  #   avg_cf_w2 = avg_cf_out2,
  #   avg_cf_w3 = avg_cf_out3,
  #   avg_cf_w4 = avg_cf_out4,
  #   avg_cf_w5 = avg_cf_out5,
  #   avg_cf_w6 = avg_cf_out6,
  #   avg_cf_w7 = avg_cf_out7,
  #   avg_cf_w8 = avg_cf_out8,
  #   avg_cf_w9 = avg_cf_out9,
  #   avg_cf_w10 = avg_cf_out10,
  #   avg_cf_w11 = avg_cf_out11,
  #   avg_cf_w12 = avg_cf_out12,
  #   avg_cf_w13 = avg_cf_out13,
  #   avg_cf_w14 = avg_cf_out14,
  #   avg_cf_w15 = avg_cf_out15,
  #   avg_cf_w16 = avg_cf_out16,
  #   avg_cf_w17 = avg_cf_out17,
  #   avg_cf_w18 = avg_cf_out18,
  #   avg_cf_w19 = avg_cf_out19,
  #   avg_cf_w20 = avg_cf_out20, 
  #   avg_cf_w21 = avg_cf_out21)
  # row.names(cf_Y_perform) = names(clustering_methods)
  # 
  # 
  # 
  # non_dominant_perform = data.frame(
  #   non_dominant_w1 = non_dominant_out1,
  #   non_dominant_w2 = non_dominant_out2,
  #   non_dominant_w3 = non_dominant_out3,
  #   non_dominant_w4 = non_dominant_out4,
  #   non_dominant_w5 = non_dominant_out5,
  #   non_dominant_w6 = non_dominant_out6,
  #   non_dominant_w7 = non_dominant_out7,
  #   non_dominant_w8 = non_dominant_out8,
  #   non_dominant_w9 = non_dominant_out9,
  #   non_dominant_w10 = non_dominant_out10,
  #   non_dominant_w11 = non_dominant_out11,
  #   non_dominant_w12 = non_dominant_out12,
  #   non_dominant_w13 = non_dominant_out13,
  #   non_dominant_w14 = non_dominant_out14,
  #   non_dominant_w15 = non_dominant_out15,
  #   non_dominant_w16 = non_dominant_out16,
  #   non_dominant_w17 = non_dominant_out17,
  #   non_dominant_w18 = non_dominant_out18,
  #   non_dominant_w19 = non_dominant_out19,
  #   non_dominant_w20 = non_dominant_out20,
  #   non_dominant_w21 = non_dominant_out21)
  # row.names(non_dominant_perform) = names(clustering_methods)
  
  cf_Y_names = c(paste0("avg_cf_mean", 1:nrow(weights)), 
                 paste0("avg_cf_sd", 1:nrow(weights)))
  cf_Y = as.data.frame(mget(cf_Y_names))
  row.names(cf_Y) = names(clustering_methods)
  
  non_dominant_names = c(paste0("non_dominant_mean", 1:nrow(weights)), 
                         paste0("non_dominant_sd", 1:nrow(weights)))
  non_dominant = as.data.frame(mget(non_dominant_names))
  row.names(non_dominant) = names(clustering_methods)
  
  
  return(list(clust_perform = clust_perform,
              # cf_Y_perform = cf_Y_perform,
              # non_dominant_perform = non_dominant_perform,
              cf_Y = cf_Y,
              true_optimal_weighted_means = true_optimal_weighted_means,
              non_dominant = non_dominant
  ))
}

```



## performance Evaluation

### Clustering Performance


```{r}
# three_outcome_result10 = Simulation_MOTree_3(iter = 100, n.cov = 5, n.per.group = 10)
# three_outcome_result10$clust_perform
# 
# three_outcome_result25 = Simulation_MOTree_3(iter = 100, n.cov = 5, n.per.group = 25)
# three_outcome_result25$clust_perform

three_outcome_result50 = Simulation_MOTree_3(iter = 100, n.cov = 5, n.per.group = 50)

three_outcome_result100 = Simulation_MOTree_3(iter = 100, n.cov = 5, n.per.group = 100)

three_outcome_result150 = Simulation_MOTree_3(iter = 100, n.cov = 5, n.per.group = 150)
```


```{r}
three_outcome_result50$clust_perform
three_outcome_result100$clust_perform
three_outcome_result150$clust_perform
```

```{r}
save(three_outcome_result10, three_outcome_result25, file = "Three_obj_results_10_25.RData")

save(#three_outcome_result10, three_outcome_result25, 
  three_outcome_result50, three_outcome_result100, three_outcome_result150, file = "Three_obj_results.RData")
```


YS May 5 update

update the depth of MOTree to be 3 instead of 2, maybe have better results
Also update the iter = 500 instead of 100.
```{r}
three_outcome_result50_new = Simulation_MOTree_3(iter = 500, n.cov = 5, n.per.group = 50)

three_outcome_result100_new = Simulation_MOTree_3(iter = 500, n.cov = 5, n.per.group = 100)

three_outcome_result150_new = Simulation_MOTree_3(iter = 500, n.cov = 5, n.per.group = 150)
```


```{r}
three_outcome_result50_new$clust_perform
three_outcome_result100_new$clust_perform
three_outcome_result150_new$clust_perform
```

For two objective case
```{r}
NDR_summary <- function(result) {
  NDR = result$non_dominant[,c(1:11)]
  summary <- data.frame(
  row_name = rownames(NDR),
  mean      = rowMeans(NDR, na.rm = TRUE),
  min       = apply(NDR, 1, min, na.rm = TRUE),
  max       = apply(NDR, 1, max, na.rm = TRUE),
  row.names = NULL            # keep row names as an explicit column
)
  return(summary)
}

NDR_summary(result50)

NDR_summary(result100)

NDR_summary(result150)
```



```{r}
View(three_outcome_result50_new$non_dominant)
View(three_outcome_result100_new$non_dominant)
View(three_outcome_result150_new$non_dominant)
```

For three objective case
```{r}
NDR_summary3 <- function(result) {
  NDR = result$non_dominant[,c(1:66)]
  summary <- data.frame(
  row_name = rownames(NDR),
  mean      = rowMeans(NDR, na.rm = TRUE),
  min       = apply(NDR, 1, min, na.rm = TRUE),
  max       = apply(NDR, 1, max, na.rm = TRUE),
  row.names = NULL            # keep row names as an explicit column
)
  return(summary)
}

NDR_summary3(three_outcome_result50_new)

NDR_summary3(three_outcome_result100_new)

NDR_summary3(three_outcome_result150_new)
```

### Visualization


$$Y_w^{*opt} - \hat Y_w^*$$
YS 04/21 update visualization

```{r}
three_obj_long_table <- function(results_summary, fineness) {
  
  n_weight = ((1/fineness) + 1) * ((1/fineness) + 2)  / 2
  outcomes <- results_summary$cf_Y[,c(1:n_weight)]
  true_opt_means <- results_summary$true_optimal_weighted_means  
  
  weights_y1_vals <- seq(0, 1, by = fineness)
  weights_y2_vals <- seq(0, 1, by = fineness)
  weights_grid <- expand.grid(Weight_Y1 = weights_y1_vals, Weight_Y2 = weights_y2_vals) %>%
    filter(Weight_Y1 + Weight_Y2 <= 1.001) %>% # Add a small tolerance for floating point issues
    mutate(Weight_Y3 = round(1 - Weight_Y1 - Weight_Y2, 1)) %>% 
    arrange(Weight_Y1, Weight_Y2) 
  
  weight_col_names <- paste0("avg_cf_mean", 1:nrow(weights_grid))
  
  weight_lookup <- weights_grid %>%
      mutate(avg_cf_col_name = weight_col_names) %>%
      mutate(Weight_Index = 1:n()) 
  
  outcomes_df <- as.data.frame(outcomes) %>%
    rownames_to_column(var = "Method")
  
  outcomes_long <- outcomes_df %>%
    pivot_longer(cols = starts_with("avg_cf_mean"),
                 names_to = "avg_cf_col_name",
                 values_to = "Estimated_Outcome")
  
  
  true_opt_means_df <- as.data.frame(true_opt_means) %>%
    pivot_longer(cols = starts_with("Weight_"),
                 names_to = "Weight_col_name",
                 values_to = "True_Optimal_Outcome") %>%
    mutate(Weight_Index = as.numeric(gsub("Weight_", "", Weight_col_name))) %>%
    dplyr::select(Weight_Index, True_Optimal_Outcome) 
  
  
  outcomes_long <- outcomes_long %>%
      mutate(Weight_Index = as.numeric(gsub("avg_cf_mean", "", avg_cf_col_name))) %>%
      dplyr::select(Method, Weight_Index, Estimated_Outcome) 
  
  
  combined_data <- left_join(outcomes_long, weight_lookup, by = "Weight_Index")
  
  combined_data <- left_join(combined_data, true_opt_means_df, by = "Weight_Index") %>%
      mutate(Difference = Estimated_Outcome - True_Optimal_Outcome) 
  
  # View(combined_data)
  return(combined_data)
}



long_table_25 = three_obj_long_table(three_outcome_result25, 0.1)

long_table_50 = three_obj_long_table(three_outcome_result50, 0.1)

long_table_100 = three_obj_long_table(three_outcome_result100, 0.1)



long_table_50_new = three_obj_long_table(three_outcome_result50_new, 0.1)

long_table_100_new = three_obj_long_table(three_outcome_result100_new, 0.1)
```


# Fiugre Idea 1: 3D Surface Plots using plotly

display separately or together 
```{r}
plot_list_3d <- lapply(unique(long_table_25$Method), function(method_name) {
  method_data <- filter(long_table_25, Method == method_name)

  # Plotly requires data for surface plots to be in a matrix format for Z values
  # We need to cast the method_data into a matrix where rows are Weight_Y1 values,
  # columns are Weight_Y2 values, and cells are the Difference.
  # However, the weight grid is not a simple rectangular grid after filtering W1+W2 <= 1.
  # Plotly can handle scattered data directly for surface plots using 'go' (graph objects)

  p <- plot_ly(method_data, x = ~Weight_Y1, y = ~Weight_Y2, z = ~Difference,
               type = 'mesh3d', # Use mesh3d or surface. Mesh3d might be better for irregular grids.
               intensity = ~Difference, colorscale = 'Viridis', # Optional: color by difference value
               name = method_name) %>%
       layout(title = paste("Method:", method_name),
              scene = list(xaxis = list(title = "Weight on Y1"),
                          yaxis = list(title = "Weight on Y2"),
                          zaxis = list(title = "Difference (Estimated - True Optimal)")))

  return(p)
})

# To display individual plots from the list (e.g., the first method's plot)
plot_list_3d[[1]]
plot_list_3d[[2]]
plot_list_3d[[3]]
plot_list_3d[[4]]
plot_list_3d[[5]]
plot_list_3d[[6]]
plot_list_3d[[7]]

# To display all plots (using subplot) - may require adjustment based on the number of methods
# Note: Subplotting 3D plots can sometimes be tricky or slow depending on the environment
subplot(plot_list_3d, nrows = ceiling(length(plot_list_3d)/1)) # Adjust nrows as needed
```

Display in one figure with color labels
```{r}
p_all_methods <- plot_ly()

for (method_name in unique(long_table_25$Method)) {
  method_data <- filter(long_table_25, Method == method_name)

  p_all_methods <- p_all_methods %>%
    add_trace(data = method_data, x = ~Weight_Y1, y = ~Weight_Y2, z = ~Difference,
              type = 'mesh3d', # Using mesh3d, consider type='surface' if data can be reshaped to a matrix
              color = ~Method, 
              name = method_name,
              opacity = 0.9) 
}

p_all_methods <- p_all_methods %>%
  layout(title = "Difference between Estimated and Optimal Counterfactual Outcome",
         scene = list(xaxis = list(title = "Weight on Y1"),
                     yaxis = list(title = "Weight on Y2"),
                     zaxis = list(title = "Difference (Estimated - True Optimal)")))


p_all_methods
```

```{r, warning = FALSE}
# method_colors2 <- c(
#   "GMM" = "#f7766d",      
#   "HC" = "#c49a00" ,       
#   "KMeans3" = "#53b401",  
#   "KMeans4" = "#02c094",  
#   "MOTree" = "#00b6eb" ,   
#   "SOM" = "#a58aff",      
#   "Spectral" = "#F781BF"  
# )

method_colors2 <- c(
  "GMM" = "#f7766d",      
  "HC" = "#efce38",  #" c49a00",       
  "KMeans3" = "#53b401",  
  "KMeans4" = "#02c094",  
  "MOTree" = "#00b6eb",   
  "SOM" = "#a58aff",      
  "Spectral" = "#F781BF"  
)

methods_in_data <- unique(long_table_25$Method)


p_all_methods <- plot_ly()

# for (method_name in names(method_colors2)) {
for (method_name in unique(long_table_25$Method)) {

  current_color <- as.character(method_colors2[method_name])
  if (is.na(current_color)) {
      current_color <- "gray" # Default color for undefined methods
  }

  method_data <- filter(long_table_25, Method == method_name)

  p_all_methods <- p_all_methods %>%
    add_trace(
      data = method_data,
      x = ~Weight_Y1,
      y = ~Weight_Y2,
      z = ~Difference,
      type = 'mesh3d', # mesh3d is suitable for your grid data
      # color = ~Method, # Remove this line, we are setting color manually per trace
      name = method_name, # This is used for the legend
      opacity = 0.8, # Slightly increased opacity for better visibility
      color = current_color, # Use I() to treat the color as a literal value
      # Add line properties for mesh visibility (optional)
      intensity = rep(1, nrow(method_data)), # Assign a constant intensity value (e.g., 1) to all points
      colorscale = list(
        c(0, current_color), # Map intensity 0 to the color
        c(1, current_color)  # Map intensity 1 to the same color
      ),
      showscale = FALSE, 
      line = list(width = 1, color = 'darkgray')
    )
}

p_all_methods <- p_all_methods %>%
  layout(
    title = "Difference between Estimated and Optimal Counterfactual Outcome",
    scene = list(
      xaxis = list(title = "Weight on Y1"),
      yaxis = list(title = "Weight on Y2"),
      zaxis = list(title = "Difference (Estimated - True Optimal)")
    ),
    margin = list(l = 0, r = 0, b = 0, t = 40), # Adjust margins
    showlegend = TRUE # Ensure legend is shown (default is usually TRUE)
  )

p_all_methods
```

Fix color 


## New Results:


### sample size n = 8 * 25

YS Note: this result using long_table_25 is from previous simulation code, with HC, SOM, GMM, Spectral clustering all based on k = 4... Which is wrong, but have better results that the updated results long_table_50 and long_table_100.
```{r}
# Define the color mapping for each method
method_colors2 <- c(
  "GMM" = "#f7766d",      
  "HC" = "#efce38",  #" c49a00",       
  "KMeans3" = "#53b401",  
  "KMeans4" = "#02c094",   
  "KMeans" = "#02c094", 
  "MOTree" = "#00b6eb",   
  "SOM" = "#bca8fd",#"#a58aff",      
  "Spectral" = "#F781BF"  
)

long_table_25_figure = long_table_25 %>%
  filter(Method != "KMeans3") %>%
  mutate(
    Method = fct_recode(Method,
      "KMeans" = "KMeans4" 
    )
  ) %>%  
  mutate(Method = droplevels(Method))

methods_in_data <- unique(long_table_25_figure$Method)

# Initialize the plot
p_all_methods <- plot_ly()

# Loop through each method to add a trace
for (method_name in unique(long_table_25_figure$Method)) {
  # Get the color for the current method
  current_color <- method_colors2[method_name]
  if (is.na(current_color)) {
    current_color <- "gray" # Default color for undefined methods
  }

  # Filter data for the current method
  method_data <- filter(long_table_25_figure, Method == method_name)

  # Add a mesh3d trace for the current method
  p_all_methods <- p_all_methods %>%
    add_trace(
      data = method_data,
      x = ~Weight_Y1,
      y = ~Weight_Y2,
      z = ~Difference,
      type = 'mesh3d',
      name = method_name, # Used for the legend
      opacity = 0.6,
      color = current_color, # Specify the exact color for this trace
      intensity = rep(1, nrow(method_data)), # Assign a constant intensity value (e.g., 1) to all points
      colorscale = list(
        c(0, current_color), # Map intensity 0 to the color
        c(1, current_color)  # Map intensity 1 to the same color
      ),
      showscale = FALSE #, 
      # line = list(width = 1, color = 'darkgray')
    )
}

# Customize the layout
p_all_methods <- p_all_methods %>%
  layout(
    title = "Difference between Estimated and Optimal Counterfactual Outcome",
    scene = list(
      xaxis = list(title = "Weight on Y1"),
      yaxis = list(title = "Weight on Y2"),
      zaxis = list(title = "Difference (Estimated - True Optimal)"),
      aspectmode = "manual", # Set aspect mode to manual control
      aspectratio = list(x = 1, y = 1, z = 1.2)
    ),

    margin = list(l = 0, r = 0, b = 0, t = 40), # Adjust margins
    showlegend = TRUE # Ensure legend is shown
  )

# Display the plot
p_all_methods
```

### sample size n = 8 * 50
```{r}
# Define the color mapping for each method
method_colors2 <- c(
  "GMM" = "#f7766d",      
  "HC" = "#efce38",  #" c49a00",       
  "KMeans5" = "#53b401",  
  "KMeans8" = "#02c094",  
  "KMeans" = "#02c094", 
  "MOTree" = "#00b6eb",   
  "SOM" = "#bca8fd",#"#a58aff",      
  "Spectral" = "#F781BF"  
)

long_table_50_figure = long_table_50 %>%
  filter(Method != "KMeans5") %>%
  mutate(
    Method = fct_recode(Method,
      "KMeans" = "KMeans8" 
    )
  ) %>%  
  mutate(Method = droplevels(Method))

methods_in_data <- unique(long_table_50_figure$Method)

# Initialize the plot
p_all_methods <- plot_ly()

# Loop through each method to add a trace
for (method_name in unique(long_table_50_figure$Method)) {
  # Get the color for the current method
  current_color <- method_colors2[method_name]
  if (is.na(current_color)) {
    current_color <- "gray" # Default color for undefined methods
  }

  # Filter data for the current method
  method_data <- filter(long_table_50_figure, Method == method_name)

  # Add a mesh3d trace for the current method
  p_all_methods <- p_all_methods %>%
    add_trace(
      data = method_data,
      x = ~Weight_Y1,
      y = ~Weight_Y2,
      z = ~Difference,
      type = 'mesh3d',
      name = method_name, # Used for the legend
      opacity = 0.6,
      color = current_color, # Specify the exact color for this trace
      intensity = rep(1, nrow(method_data)), # Assign a constant intensity value (e.g., 1) to all points
      colorscale = list(
        c(0, current_color), # Map intensity 0 to the color
        c(1, current_color)  # Map intensity 1 to the same color
      ),
      showscale = FALSE #, 
      # line = list(width = 1, color = 'darkgray')
    )
}

# Customize the layout
p_all_methods <- p_all_methods %>%
  layout(
    title = "Difference between Estimated and Optimal Counterfactual Outcome",
    scene = list(
      xaxis = list(title = "Weight on Y1"),
      yaxis = list(title = "Weight on Y2"),
      zaxis = list(title = "Difference (Estimated - True Optimal)"),
      aspectmode = "manual", # Set aspect mode to manual control
      aspectratio = list(x = 1, y = 1, z = 1.2)
    ),

    margin = list(l = 0, r = 0, b = 0, t = 40), # Adjust margins
    showlegend = TRUE # Ensure legend is shown
  )

# Display the plot
p_all_methods
```

new
```{r}
# Define the color mapping for each method
method_colors2 <- c(
  "GMM" = "#f7766d",      
  "HC" = "#efce38",  #" c49a00",       
  "KMeans5" = "#53b401",  
  "KMeans8" = "#02c094",  
  "KMeans" = "#02c094", 
  "MOTree" = "#00b6eb",   
  "SOM" = "#bca8fd",#"#a58aff",      
  "Spectral" = "#F781BF"  
)

long_table_50_figure = long_table_50_new %>%
  filter(Method != "KMeans5") %>%
  mutate(
    Method = fct_recode(Method,
      "KMeans" = "KMeans8" 
    )
  ) %>%  
  mutate(Method = droplevels(Method))

methods_in_data <- unique(long_table_50_figure$Method)

# Initialize the plot
p_all_methods <- plot_ly()

# Loop through each method to add a trace
for (method_name in unique(long_table_50_figure$Method)) {
  # Get the color for the current method
  current_color <- method_colors2[method_name]
  if (is.na(current_color)) {
    current_color <- "gray" # Default color for undefined methods
  }

  # Filter data for the current method
  method_data <- filter(long_table_50_figure, Method == method_name)

  # Add a mesh3d trace for the current method
  p_all_methods <- p_all_methods %>%
    add_trace(
      data = method_data,
      x = ~Weight_Y1,
      y = ~Weight_Y2,
      z = ~Difference,
      type = 'mesh3d',
      name = method_name, # Used for the legend
      opacity = 0.6,
      color = current_color, # Specify the exact color for this trace
      intensity = rep(1, nrow(method_data)), # Assign a constant intensity value (e.g., 1) to all points
      colorscale = list(
        c(0, current_color), # Map intensity 0 to the color
        c(1, current_color)  # Map intensity 1 to the same color
      ),
      showscale = FALSE #, 
      # line = list(width = 1, color = 'darkgray')
    )
}

# Customize the layout
p_all_methods <- p_all_methods %>%
  layout(
    title = "Difference between Estimated and Optimal Counterfactual Outcome",
    scene = list(
      xaxis = list(title = "Weight on Y1"),
      yaxis = list(title = "Weight on Y2"),
      zaxis = list(title = "Difference (Estimated - True Optimal)"),
      aspectmode = "manual", # Set aspect mode to manual control
      aspectratio = list(x = 1, y = 1, z = 1.2)
    ),

    margin = list(l = 0, r = 0, b = 0, t = 40), # Adjust margins
    showlegend = TRUE # Ensure legend is shown
  )

# Display the plot
p_all_methods
```

### sample size n = 8 * 100
```{r}
# Define the color mapping for each method
method_colors2 <- c(
  "GMM" = "#f7766d",      
  "HC" = "#efce38",  #" c49a00",       
  "KMeans3" = "#53b401",  
  "KMeans4" = "#02c094",  
  "MOTree" = "#00b6eb",   
  "SOM" = "#bca8fd",#"#a58aff",      
  "Spectral" = "#F781BF"  
)

# Get unique methods from the data
methods_in_data <- unique(long_table_100$Method)

# Initialize the plot
p_all_methods <- plot_ly()

# Loop through each method to add a trace
for (method_name in unique(long_table_100$Method)) {
  # Get the color for the current method
  current_color <- method_colors2[method_name]
  if (is.na(current_color)) {
    current_color <- "gray" # Default color for undefined methods
  }

  # Filter data for the current method
  method_data <- filter(long_table_100, Method == method_name)

  # Add a mesh3d trace for the current method
  p_all_methods <- p_all_methods %>%
    add_trace(
      data = method_data,
      x = ~Weight_Y1,
      y = ~Weight_Y2,
      z = ~Difference,
      type = 'mesh3d',
      name = method_name, # Used for the legend
      opacity = 0.6,
      color = current_color, # Specify the exact color for this trace
      intensity = rep(1, nrow(method_data)), # Assign a constant intensity value (e.g., 1) to all points
      colorscale = list(
        c(0, current_color), # Map intensity 0 to the color
        c(1, current_color)  # Map intensity 1 to the same color
      ),
      showscale = FALSE #, 
      # line = list(width = 1, color = 'darkgray')
    )
}

# Customize the layout
p_all_methods <- p_all_methods %>%
  layout(
    title = "Difference between Estimated and Optimal Counterfactual Outcome",
    scene = list(
      xaxis = list(title = "Weight on Y1"),
      yaxis = list(title = "Weight on Y2"),
      zaxis = list(title = "Difference (Estimated - True Optimal)"),
      aspectmode = "manual", # Set aspect mode to manual control
      aspectratio = list(x = 1, y = 1, z = 1.2)
    ),

    margin = list(l = 0, r = 0, b = 0, t = 40), # Adjust margins
    showlegend = TRUE # Ensure legend is shown
  )

# Display the plot
p_all_methods
```

new
```{r}
# Define the color mapping for each method
method_colors2 <- c(
  "GMM" = "#f7766d",      
  "HC" = "#efce38",  #" c49a00",       
  "KMeans3" = "#53b401",  
  "KMeans4" = "#02c094",  
  "MOTree" = "#00b6eb",   
  "SOM" = "#bca8fd",#"#a58aff",      
  "Spectral" = "#F781BF"  
)

# Get unique methods from the data
methods_in_data <- unique(long_table_100_new$Method)

# Initialize the plot
p_all_methods <- plot_ly()

# Loop through each method to add a trace
for (method_name in unique(long_table_100_new$Method)) {
  # Get the color for the current method
  current_color <- method_colors2[method_name]
  if (is.na(current_color)) {
    current_color <- "gray" # Default color for undefined methods
  }

  # Filter data for the current method
  method_data <- filter(long_table_100_new, Method == method_name)

  # Add a mesh3d trace for the current method
  p_all_methods <- p_all_methods %>%
    add_trace(
      data = method_data,
      x = ~Weight_Y1,
      y = ~Weight_Y2,
      z = ~Difference,
      type = 'mesh3d',
      name = method_name, # Used for the legend
      opacity = 0.6,
      color = current_color, # Specify the exact color for this trace
      intensity = rep(1, nrow(method_data)), # Assign a constant intensity value (e.g., 1) to all points
      colorscale = list(
        c(0, current_color), # Map intensity 0 to the color
        c(1, current_color)  # Map intensity 1 to the same color
      ),
      showscale = FALSE #, 
      # line = list(width = 1, color = 'darkgray')
    )
}

# Customize the layout
p_all_methods <- p_all_methods %>%
  layout(
    title = "Difference between Estimated and Optimal Counterfactual Outcome",
    scene = list(
      xaxis = list(title = "Weight on Y1"),
      yaxis = list(title = "Weight on Y2"),
      zaxis = list(title = "Difference (Estimated - True Optimal)"),
      aspectmode = "manual", # Set aspect mode to manual control
      aspectratio = list(x = 1, y = 1, z = 1.2)
    ),

    margin = list(l = 0, r = 0, b = 0, t = 40), # Adjust margins
    showlegend = TRUE # Ensure legend is shown
  )

# Display the plot
p_all_methods
```






# --- Discussion on Visualization Methods ---

(1) 3D Surface Plot with Smoothing:
Pros: Directly visualizes the performance metric (difference) across the weight space. Surfaces provide a sense of continuity. Plotly allows for interactive exploration.
Cons: Surfaces can overlap, making individual method performance hard to distinguish, especially with many methods. The grid is not perfectly rectangular, so `mesh3d` is often better than `surface` for scattered data like this, but smoothing isn't automatic. You might need to use smoothing techniques *before* plotting or explore Plotly's surface options carefully.

(2) 3D Histogram:
Pros: Can show the *distribution* of the difference values for specific weight bins. Might be useful for discrete weight points if you have many iterations per weight.
Cons: Doesn't naturally form a continuous surface across the weight space. A single figure with multiple 3D histograms side-by-side would be cluttered. Separate figures for each method would lose the direct comparison benefit. It's less intuitive for showing performance *across* the continuous weight spectrum compared to surfaces. Generally not ideal for showing function approximation performance over a continuous domain.

(3) Sphere-like figure setting with 3-axis coordinate system:
Concept: If you mean plotting the *estimated outcomes* (Y_hat_w1, Y_hat_w2, Y_hat_w3) in a 3D space colored by method or weight, this would show the *predicted* Pareto frontier approximation. Plotting the *difference* (Y_hat - Y_opt) in a 3D outcome space doesn't seem intuitive, as the difference is a scalar value for each weight vector. The weight vector itself lives in a different space (the unit simplex).
Pros: If plotting the estimated outcomes, it shows how well methods approximate the true Pareto front in the outcome space.
Cons: Doesn't directly visualize the performance (difference) across the weight space. Visualizing the *difference* in the outcome space is conceptually less clear. The "sphere-like" idea might be related to visualizing points on the unit sphere for directions in outcome space, but your weights are on the *input* side.

(4) Drawing lines on planes (Y1=0, Y2=0, Y3=0):
Concept: This visualizes cross-sections of the 3D surface from idea (1).
For the Y1=0 plane, you'd plot Difference vs. Weight_Y2 (where Weight_Y1=0, so Weight_Y3 = 1 - Weight_Y2).
For the Y2=0 plane, you'd plot Difference vs. Weight_Y1 (where Weight_Y2=0, so Weight_Y3 = 1 - Weight_Y1).
For the Y3=0 plane, you'd plot Difference vs. Weight_Y1 (where Weight_Y3=0, so Weight_Y2 = 1 - Weight_Y1).
Pros: Breaks down the 3D visualization into more manageable 2D plots. Allows direct line-based comparison on specific "slices" of the weight space (e.g., when one weight is zero). Easier to interpret than overlapping 3D surfaces.
Cons: Only shows performance on the boundaries of the weight space (when one weight is zero), not the internal space.


# Figure Idear 2: Lines on Planes 

This requires filtering the data for each plane and plotting in 2D.

```{r}
plane_y1_0_data <- long_table_25 %>% filter(Weight_Y1 == 0)

plot_y1_0 <- ggplot(plane_y1_0_data, aes(x = Weight_Y2, y = Difference, color = Method)) +
  geom_line(size = 1.2) + # geom_smooth() +
  labs(title = "Performance when Weight on Y1 = 0",
       x = "Weight on Y2",
       y = "Difference (Estimated - True Optimal)") +
  theme_minimal()

print(plot_y1_0)

plane_y2_0_data <- long_table_25 %>% filter(Weight_Y2 == 0)

plot_y2_0 <- ggplot(plane_y2_0_data, aes(x = Weight_Y1, y = Difference, color = Method)) +
  geom_line(size = 1.2) + # Or geom_smooth
  labs(title = "Performance when Weight on Y2 = 0",
       x = "Weight on Y1",
       y = "Difference (Estimated - True Optimal)") +
  theme_minimal()

print(plot_y2_0)

plane_y3_0_data <- long_table_25 %>% filter(Weight_Y3 == 0)

plot_y3_0 <- ggplot(plane_y3_0_data, aes(x = Weight_Y1, y = Difference, color = Method)) +
  geom_line(size = 1.2) + # Or geom_smooth
  labs(title = "Performance when Weight on Y3 = 0",
       x = "Weight on Y1",
       y = "Difference (Estimated - True Optimal)") +
  theme_minimal()

print(plot_y3_0)
```


# Figure Idear 3: Contour Plot on Ternary Diagram

```{r}
# Concept: Ternary diagrams are designed specifically for visualizing data on a 3-component system where the components sum to 1 (like your weights W1, W2, W3). You can plot contour lines or a heatmap of the 'Difference' value on this 2D triangular plane.
# Pros: Directly represents the entire 3D weight space in a single 2D plot. Intuitive for weight combinations. Can show how performance changes as you move between focusing on different outcomes.
# Cons: Requires a specialized library for ternary plots. Plotting multiple methods on the same ternary plot might lead to clutter unless using faceting or careful layering.

# R Package Needed: `ggtern` or `ternary`

# R Code (using ggtern - might need installation: install.packages("ggtern"))
library(ggtern)
library(Ternary)

# # ggtern requires data in a specific format and column names W1, W2, W3
ternary_data <- long_table_25 %>%
  rename(W1 = Weight_Y1, W2 = Weight_Y2, W3 = Weight_Y3)

# # Create a ternary contour plot faceted by method
plot_ternary_contour <- ggtern(data = ternary_data, aes(x = W1, y = W2, z = W3, value = Difference)) +
  geom_interpolate_contour(aes(color = ..value..)) + # Contour lines based on difference
  # geom_density_tern() + # Another option - density estimate if you had point data not grid
  facet_wrap(~ Method) + # Create a separate plot for each method
  labs(title = "Performance Difference on Ternary Weight Space",
       fill = "Difference") + # Label for the color scale
  theme_bw() +
  theme_tern_common()

print(plot_ternary_contour)

# # Or a ternary heatmap (tile plot)
plot_ternary_tile <- ggtern(data = ternary_data, aes(x = W1, y = W2, z = W3, fill = Difference)) +
  geom_point() + # Plot the data points
  # Consider using geom_interpolate_fill() if points are on a grid, or geom_confidence()
  facet_wrap(~ Method) +
  scale_fill_gradient2() + # Use a diverging color scale for differences around 0
  labs(title = "Performance Difference on Ternary Weight Space",
       fill = "Difference") +
  theme_bw() +
  theme_tern_common()

print(plot_ternary_tile)
```

updated
```{r}
ternary_data <- long_table_25 %>%
  rename(W1 = Weight_Y1, W2 = Weight_Y2, W3 = Weight_Y3)

plot_ternary_contour <- ggtern(data = ternary_data, aes(x = W1, y = W2, z = W3, value = Difference)) +
  stat_interpolate_tern(aes(color = after_stat(value), group = Method), geom = "contour") + # Use stat_interpolate_tern + geom_contour
  # The 'value' aesthetic in stat_interpolate_tern is the variable to interpolate (Difference).
  # 'after_stat(value)' refers to the interpolated value produced by the stat.
  # 'group = Method' ensures interpolation is done separately for each method.
  # Optional: Add the actual points if you want to see the data locations
  geom_point(size = 0.5, alpha = 0.5) +
  facet_wrap(~ Method) + # Create a separate plot for each method
  scale_color_viridis_c(name = "Difference") + # Use a color scale for the contour lines
  labs(title = "Performance Difference on Ternary Weight Space (Contour Lines)") +
  theme_bw()  +
  theme_tern()

print(plot_ternary_contour)


# Create a ternary filled contour / heatmap faceted by method
# Using stat_interpolate_tern with geom_polygon to create filled areas
plot_ternary_fill <- ggtern(data = ternary_data, aes(x = W1, y = W2, z = W3, value = Difference)) +
  stat_interpolate_tern(aes(fill = after_stat(value), group = Method), geom = "polygon") + # Use stat_interpolate_tern + geom_polygon
  # 'after_stat(value)' refers to the interpolated value for coloring the polygons.
  # 'group = Method' ensures interpolation is done separately for each method.
  # Optional: Add the actual points if you want to see the data locations
  # geom_point(size = 0.5, alpha = 0.5) +
  facet_wrap(~ Method) +
  scale_fill_gradient2(name = "Difference") + # Use a diverging color scale for differences around 0
  labs(title = "Performance Difference on Ternary Weight Space (Filled Contours)") +
  theme_bw() +
  theme_tern()

print(plot_ternary_fill)
```



# --- Recommendation ---
The most direct visualization of the difference (Y_hat - Y_opt) across the 2D weight simplex is either:

1.  **3D Surface Plot (Option 1):** This gives a sense of the landscape of performance across the weight space. Using `plotly` with `mesh3d` and opacity is a good approach to show all methods in one interactive figure, even if surfaces overlap.
2.  **Ternary Contour/Heatmap (Option 5):** This collapses the 3D weight space into a 2D triangle, which is a standard way to visualize 3-component compositions. Contour lines or color gradients clearly show how performance changes across the simplex. Faceting by method allows for clear comparison.

Option (4) (lines on planes) is also valuable as it provides easily interpretable 2D slices, highlighting performance on the boundaries of the weight space. It complements the 3D or ternary views.

Option (2) (3D histogram) is generally not suitable for visualizing continuous performance over a weight spectrum. Option (3) (sphere-like outcome plot) visualizes the predicted outcome space rather than the performance difference across the input weight space.

I recommend starting with the `plotly` 3D surface plot (Option 1) for an interactive overview and the ternary plot (Option 5) or the 2D plane plots (Option 4) for more detailed, static views. The ternary plot is arguably the most theoretically appropriate for visualizing data tied to a compositional (weight) space.

The R code provided creates the long data frame `combined_data` and then shows how to generate the 3D surface plots for all methods in one figure using `plotly`. It also includes the code for the 2D plane plots as an alternative. Code for ternary plots is commented out as it requires an additional library.

```{r}
# generating a long table outcomes_long_3d


# Create a 3D surface plot for each method
plot_list <- lapply(unique(outcomes_long_3d$Method), function(method) {
  method_data <- filter(outcomes_long_3d, Method == method)
  plot_ly(method_data, x = ~Weight_Y1, y = ~Weight_Y2, z = ~Difference, type = "surface") %>%
    layout(title = paste("Method:", method),
           scene = list(xaxis = list(title = "Weight on Y1"),
                       yaxis = list(title = "Weight on Y2"),
                       zaxis = list(title = "Difference (Y_opt - Y_hat)")))
})

# For individual plots
# i <- 1  # Replace with the index of the method you want to view (1 to number of methods)
# plot_list[[i]]

# To display all plots (using `subplot`)
subplot(plot_list, nrows = 2, shareX = TRUE, shareY = TRUE)


```


# Archive

```{r}
# Extract outcomes from the simulation results
outcomes <- result50$cf_Y  # Counterfactual outcomes
true_opt_means <- result50$true_optimal_weighted_means  # 1x11 matrix of observed weighted outcomes

# Prepare the observed means for merging
true_opt_long <- as.data.frame(t(true_opt_means)) %>%
  rownames_to_column(var = "Weight") %>%
  rename(TrueOpt = True_Optimal_Mean) %>%
  mutate(Weight = as.numeric(gsub("Weight_", "", Weight)))

# Prepare the counterfactual mean outcomes and merge with observed means
outcomes_long <- outcomes %>%
  rownames_to_column(var = "Method") %>%
  pivot_longer(cols = starts_with("avg_cf_mean"), 
               names_to = "Weight", 
               values_to = "Mean") %>%
  mutate(Weight = as.numeric(gsub("avg_cf_mean", "", Weight)) * 0.1) %>%
  left_join(true_opt_long, by = "Weight") %>%
  mutate(Difference = Mean - TrueOpt)  # Calculate the difference between counterfactual and observed means

# Plot the differences
ggplot(outcomes_long, aes(x = Weight, y = Difference, color = Method, fill = Method)) +
  geom_smooth(method = "loess", se = TRUE, size = 1.2) +  # Smoothed line with CI for difference
  labs(title = "Difference between Estimated and True Optimal Conterfactual Outcome",
       x = "Weight on Y1",
       y = "Difference (Y_opt - Y_hat)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.title = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )
```

```{r}
# Extract the non-dominant data from the simulation results
non_dominant <- result50$non_dominant

# Prepare the non-dominant data for plotting
non_dominant_long <- non_dominant %>%
  rownames_to_column(var = "Method") %>%
  pivot_longer(cols = starts_with("non_dominant_mean"), 
               names_to = "Weight", 
               values_to = "Mean") %>%
  mutate(Weight = as.numeric(gsub("non_dominant_mean", "", Weight)) * 0.1) %>%
  left_join(
    non_dominant %>%
      rownames_to_column(var = "Method") %>%
      pivot_longer(cols = starts_with("non_dominant_sd"), 
                   names_to = "Weight", 
                   values_to = "SD") %>%
      mutate(Weight = as.numeric(gsub("non_dominant_sd", "", Weight)) * 0.1),
    by = c("Method", "Weight")
  )

# Plot the percentage of non-dominance
ggplot(non_dominant_long, aes(x = Weight, y = Mean, color = Method, fill = Method)) +
  geom_line(size = 1.2) +  # Line for the mean percentage of non-dominance
  geom_ribbon(aes(ymin = Mean - 1.96*SD/sqrt(200), ymax = Mean + 1.96*SD/sqrt(200)), alpha = 0.2, linetype = 0) +  # Add shadow for SD
  labs(title = "Percentage of Non-Dominance of the Estimated Optimal Treatment",
       x = "Weight on Y1",
       y = "Percentage of Non-Dominance") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.title = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )
```




Sample size = 4 * 500
```{r}
# Extract outcomes from the simulation results
outcomes <- result500$cf_Y  # Counterfactual outcomes
true_opt_means <- result500$true_optimal_weighted_means  # 1x11 matrix of observed weighted outcomes

# Prepare the observed means for merging
true_opt_long <- as.data.frame(t(true_opt_means)) %>%
  rownames_to_column(var = "Weight") %>%
  rename(TrueOpt = True_Optimal_Mean) %>%
  mutate(Weight = as.numeric(gsub("Weight_", "", Weight)))

# Prepare the counterfactual mean outcomes and merge with observed means
outcomes_long <- outcomes %>%
  rownames_to_column(var = "Method") %>%
  pivot_longer(cols = starts_with("avg_cf_mean"), 
               names_to = "Weight", 
               values_to = "Mean") %>%
  mutate(Weight = as.numeric(gsub("avg_cf_mean", "", Weight)) * 0.1) %>%
  left_join(true_opt_long, by = "Weight") %>%
  mutate(Difference = Mean - TrueOpt)  # Calculate the difference between counterfactual and observed means

# Plot the differences
ggplot(outcomes_long, aes(x = Weight, y = Difference, color = Method, fill = Method)) +
  geom_smooth(method = "loess", se = TRUE, size = 1.2) +  # Smoothed line with CI for difference
  labs(title = "Difference between Estimated and True Optimal Conterfactual Outcome",
       x = "Weight on Y1",
       y = "Difference (Y_opt - Y_hat)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.title = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )


```

```{r}
# Extract the non-dominant data from the simulation results
non_dominant <- result500$non_dominant

# Prepare the non-dominant data for plotting
non_dominant_long <- non_dominant %>%
  rownames_to_column(var = "Method") %>%
  pivot_longer(cols = starts_with("non_dominant_mean"), 
               names_to = "Weight", 
               values_to = "Mean") %>%
  mutate(Weight = as.numeric(gsub("non_dominant_mean", "", Weight)) * 0.1) %>%
  left_join(
    non_dominant %>%
      rownames_to_column(var = "Method") %>%
      pivot_longer(cols = starts_with("non_dominant_sd"), 
                   names_to = "Weight", 
                   values_to = "SD") %>%
      mutate(Weight = as.numeric(gsub("non_dominant_sd", "", Weight)) * 0.1),
    by = c("Method", "Weight")
  )

# Plot the percentage of non-dominance
ggplot(non_dominant_long, aes(x = Weight, y = Mean, color = Method, fill = Method)) +
  geom_line(size = 1.2) +  # Line for the mean percentage of non-dominance
  geom_ribbon(aes(ymin = Mean - 1.96*SD/sqrt(200), ymax = Mean + 1.96*SD/sqrt(200)), alpha = 0.2, linetype = 0) +  # Add shadow for SD
  labs(title = "Percentage of Non-Dominance of the Estimated Optimal Treatment",
       x = "Weight on Y1",
       y = "Percentage of Non-Dominance") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
    axis.title = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )

```



### 12/02/2024 YS to do:

Three outcome scenario

- [Table for clustering results] 

- [Figure for counterfactrual outcome and dominance]


**Add more scenarios**

- [add to Table 1 as new chunks] design other scenarios with different sample size? noise? covaraite structure
Sample sizes: 50, 100, 200
Current noise: Normal(0, 0.3)
Current covaraites: 5



